Self-assembly of colloid-cholesteric composites: A route to switchable optical materials

### 2013-10-02 papers/2012-sandwich/support.tex
Updates to supporting material
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2156 40876441-3496-0410-a613-a2fca35b2ae2

Explicitly, the free energy density is:
\begin{eqnarray}
f(Q_{\alpha\beta}) &=& {\textstyle \frac{1}{2}} A_0
 \left(1- {\textstyle \frac{1}{3}}\gamma \right)Q^2_{\alpha \beta}
-{\textstyle \frac{1}{3}}A_0\gamma Q_{\alpha \beta} Q_{\beta \gamma}Q_{\gamma \alpha} 
+ {\textstyle \frac{1}{4}} A_0\gamma (Q^2_{\alpha \beta})^2  \nonumber\\
&+& {\textstyle \frac{1}{2}} K(\varepsilon_{\alpha \gamma \delta}
\partial_\gamma Q_{\delta \beta} + 2 q_0 Q_{\alpha \beta})^2
+ {\textstyle \frac{1}{2}} K (\partial_\beta Q_{\alpha \beta})^2
\label{free}
\end{eqnarray}
Here, repeated indices are summed over, while terms of the form
$Q^2_{\alpha\beta}$ should be expanded to read $Q_{\alpha\beta}Q_{\alpha\beta}$;
$\varepsilon_{\alpha\gamma\delta}$ is the permutation tensor.
The first three terms are a bulk free energy density whose overall scale
is set by $A_0$ (discussed further below); $\gamma$ is a control parameter
related to a reduced temperature. Varying the latter in the absence of
chiral terms ($q_0=0$) gives an isotropic-nematic transition at
$\gamma = 2.7$ with a mean-field spinodal instability at $\gamma = 3$.

The remaining terms of the free energy density in Eq.~\ref{free} describe
distortions of the order parameter field. In theoretical work and when
describing a generic rather than a specific system, it is 
conventional~\cite{blue1,deGennes} to
assume that splay, bend and twist deformations of the director are equally
costly, that is, there is a single elastic constant $K$. The parameter $q_0$
is related to the helical pitch length $p$ via $q_0=2\pi/p$,
describing one full turn of the director in the cholesteric phase.

There are two dimensionless numbers, which are commonly referred to as
$\kappa$, the chirality, and  $\tau$, the reduced temperature
\cite{blue1}.  In terms of the parameters in the free energy density,
these are:
\begin{eqnarray}\label{cntrl-param} 
\tau&=&\frac{27(1-\gamma/3)}{\gamma}\label{tau}\\
\kappa&=&\sqrt{\frac{108\ K\, q_0^2}{A_0\, \gamma}}\label{kappa}.
\end{eqnarray}
If the free energy density Eq.~\ref{free} is made dimensionless, $\tau$
appears as prefactor of the term  quadratic in $Q_{\alpha\beta}$,
whereas $\kappa$
quantifies the ratio between bulk and  gradient free energy terms. The
chirality and reduced temperature may be used to characterise an equilibrium
phase diagram of the blue phases in bulk cholesteric liquid crystals
\cite{blue1,oliver1}.
In this work, the chirality and reduced temperature are always set to lie
within the blue phase I region of the equilibrium phase diagram.

@@

The effect of an electric field can be modelled by including the following
term in the free energy density,
\begin{equation}
-(\epsilon_a/12\pi)E_{\alpha}Q_{\alpha\beta}E_{\beta}
\end{equation} 
where $E_{\beta}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
a further dimensionless number, 
\begin{equation}
{\cal E}^2=(27 \varepsilon_a E_\alpha E_\alpha)/(32 \pi A_0 \gamma).
\end{equation}

@@

The confining solid walls, which are flat and stationary, have the
same normal boundary condition as at colloid surface. In addition,
degenerate planar anchoring is applied at the walls, where the
preferred nematic director may lay in any orientation parallel
to the surface. The preferred direction is determined explicitly
by projecting the local fluid order parameter to the plane of the
wall \cite{fournier2005}. Eq.~\ref{eq:surf_q} and Eq.~\ref{eq:surf_q0}
are then used as before.

@@

The parameters for the free energy are chosen to be representative
of blue phase I: chirality $\kappa = 0.7$ and reduced temperature
$\tau = 0$.
The full free energy parameters are shown in Table~\ref{table:params}.
The order parameter tensor $Q_{\alpha\beta}$ for bulk blue phase I is
initialised from an approximation in high chirality limit
\cite{blue1,oliver1}. Systems of 128$^3$ lattice sites are
used with a pitch length of $p = 32\sqrt{2}$, which accommodates 4$^3$
unit cells with fully periodic boundary conditions.
Simulations at different solid volume fractions (1\% and 5\%), and different 
surface anchoring strengths ($Wa/K = 0.23$ and $Wa/k = 2.3$) representing
``weak'' and ``strong'' anchoring are run for 500,000 simulation
time steps (and 700,000 time steps for 5\% solid volume fraction).
The initial distribution of the colloid positions is set at random
for the required solid volume fraction; the colloids are initially
at rest.

To generate a disordered network of disclination lines, a ``quench''
is performed in which the order parameter is initialised via a
locally chosen random director (and using an equation of the form
Eq.~\ref{eq:surf_q} with an initial amplitude of order $q = -0.2$).
In order to avoid the mean-field spinodal point at $\gamma = 3.0$,
the parameters are adjusted slightly while maintaining the same
pitch $p = 32\sqrt{2}$. The chirality and reduced temperature (see
Table~\ref{table:params}) remain appropriate for blue phase I.
Colloids are added as before and the simulations run for 500,000 or 
700,000 simulation steps.

@@

Here, a narrow sandwich of fluid is placed between flat walls in
perpendicular to the narrow coordinate direction, and with periodic
boundaries in the other two directions. The system size used in all
cases is 256$^2 \times$56 lattice sites. Each simulation is a ``quench''
in a similar manner to that used in the bulk: the order parameter is
initialised randomly. The free energy parameters (see Table~\ref{table:params})
are again appropriate for blue phase I (chirality $\kappa = 0.8$ and
reduced temperature $\tau = -0.25$). The cholesteric pitch is set
to be $p = 64$, providing a slight better resolution that the bulk
simulations.

The surface anchoring for colloids is as before: always normal, but with
$Wa/K = 0.23$ and $Wa/K = 2.3$. The colloid solid volume fractions are
1\% and 4\%. Wall anchoring is either normal or planar with $W/K = 1$
in simulation units. All simulations are run for 1.6 million simulations
time steps unless otherwise stated CHECK.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
         & $A_0$ & $\gamma$ & $K$ & $p$ & $\kappa$ & $\tau$ & $Wa/K$\\
\hline 
Bulk BPI & 0.01 & 3.000 & 0.007061 & 32$\sqrt{2}$ & 0.7000 & 0.0000 &
0.23, 2.3\\
\hline
Bulk Quench & 0.01 & 3.100 & 0.007061 & 32$\sqrt{2}$ & 0.6886 & -0.2903 &
0.23, 2.3 \\
\hline
Confined & 0.01 & 3.0857 & 0.01897 & 64 & 0.8 & -0.25 & 0.23, 2.3\\
\hline
\end{tabular}
\end{center}
\caption{Free energy parameters used in the simulations.}
\label{table:params}
\end{table}

In all cases visualisation of the defect lines is carried out via
an isosurface of the scalar order parameter $q$ determined from the
largest eigenvalue of the local order parameter tensor. A low value
of $q$ (typically 0.12) unambiguously identifies the disclinations.

@@

To get from simulation to physical units, we need to calibrate scales of length,
energy and time. For similar mappings, see e.g.~\cite{denniston2}.

The length scale can be set by mapping our half pitch to realistic blue phase
unit cell size, which are in the 100-500 nm range \cite{blue1}. 
For bulk simulations (Fig. 1 and Fig. 3 in the main text), a suitable choice
is that that one simulation unit (lattice site) corresponds to 10 nm. Therefore
the colloidal size in Figs. 1 and 3 corresponds to $\sim$ 50 nm.

To get an energy scale, we choose $A_0 \simeq 10^6$ Pa, which is 
reasonable from Ref.~\cite{blue1} (this is also the choice in
Ref.~\cite{oliver2}). This leads to a simulation unit of
free energy (or stress) equal to $10^{8}$Pa in Figs. 1 and 3.
From the energy and length scales one can map elastic constants to
physical units: for instance the simulation of Fig. 1 corresponds to
a liquid crystal with splay, bend and twist (Frank) elastic constants
equal to 35 pN (or equivalently $K=$70 pN).
\begin{equation}
\frac{a}{b^2} = \frac{27}{2 A_0\gamma} \sim 2-5 \times 10^{-6}
{\rm J}^{-1}\, {\rm m}^{3}.
\end{equation}

For the timescale calibration we use the formula (see e.g.~\cite{denniston1})
\begin{equation}
\gamma_1=\frac{2q^2}{\Gamma}\;.
\end{equation}
 which relates the `rotational viscosity' $\gamma_1$, to the ordering strength $q$ and the order parameter mobility $\Gamma$.
In our simulations, we chose $\Gamma = 0.5$, while the value of $\gamma$ used in Fig. 1 leads to $q=1/2$.   
For real liquid crystalline materials, $\gamma_1$ lies usually in range $10^{-2}-1$ Pa s~\cite{deGennes}; we choose for definiteness $\gamma_1 = 1$ Pa s (equivalently 10 poise). Given the previous mapping for free energy (or stress) units, this leads to a simulation unit of time equal to $10^{-8}$ s. The total simulated time in Figs. 1-3 is therefore in the ms range.

To calibrate the electric field strength 
we equate the value of the dimensionless parameter ${\cal E}$ in our 
simulations with that of a hypothetical experiment.
Given the previous value for $A_0$, and a dielectric anisotropy of 10
(or equivalently $\epsilon_a=10\epsilon_0$, with $\epsilon_0$ the
dielectric permittivity of vacuum), we then obtain that an electric field
of 5 V/$\mu$m corresponds to a value of ${\cal E}\sim$0.01.

Given these mappings, all quantities can be easily transformed from
simulation to physical units and vice versa. For example Fig. 1 and 2
use an isotropic viscosity $\eta \simeq$ 0.01 and 0.1 in simulation units,
which translate into $0.01-0.1$ Pa s respectively. This is sensible (if 
slightly low) for a molecular nematogen in the isotropic phase. 
The effective viscosity in ordered phases is higher, but this is ensured
in our model by the coupling to the order parameter.
We note that the density is set to 1 in our simulations: this
corresponds to a fluid density which is about a thousand times larger
than in experiments. This however makes no difference in practice as
the Reynolds number $= \rho V\Lambda/\eta$ (where $V$ is a typical
velocity, in our simulations around $10^{-5}$ in simulation units) 
remains small enough~\cite{codef}.

To summarize the above, our simulations represent BP-forming materials, with the interpretation of the LB units for length, time, and energy density are close to 10 nm, 10 ns, and 100 MPa respectively.

@@

%ISSUE: no redshift?

@@

{\bf 6. Supporting Figures and Animations}

\begin{figure}
\begin{center}
\includegraphics[scale=0.35]{s1.jpg}
\end{center}
\caption{As for main manuscript Fig.~1A-E (initialised to equilibrium blue
phase I), but for the entire simulation system of 128$^3$ lattice sites. (a) 
solid volume fraction 1\% and weak anchoring; 
(b) 1\% solid volume fraction and strong anchoring;
(c) 5\% solid volume fraction and weak anchoring; (d) 5\% solid
volume fraction and strong anchoring. The view direction in the
main figure is from the left here. There is a reference
particle at the bottom left in each panel which does not take part
in the simulation.}
\end{figure}

\begin{figure}
\includegraphics[scale=0.35]{s2.jpg}
\caption{As for main manuscript Fig.~1E-H (initialised via a ``quench''
to generate a disordered network), but for the entire simulation
system of 128$^3$ lattice sites. (a) solid volume fraction 1\% and
weak anchoring; (b) 1\% solid volume fraction and strong anchoring;
(c) 5\% solid volume fraction and weak anchoring; (d) 5\% solid
volume fraction and strong anchoring. Again, there is a reference
particle at the bottom left in each case which does not take part
in the simulation.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.42]{s3.jpg}
\end{center}
\caption{As for main manuscript Fig.~3 (confined geometry with normal
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (a) solid volume fraction 1\%
and weak anchoring; (b) 1\% solid volume fraction and strong anchoring;
(c) 4\% solid volume fraction and weak anchoring; (d) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.42]{s4.jpg}
\end{center}
\caption{As for main manuscript Fig.~4 (confined geometry with planar
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (a) colloid solid volume fraction 1\%
and weak anchoring; (b) 1\% solid volume fraction and strong anchoring;
(c) 4\% solid volume fraction and weak anchoring; (d) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.95\columnwidth]{s5.jpg}
\end{center}
\caption{Pair correlation function for weak ($WR/K=0.23$) and strong ($WR/K=2.3$) anchoring cases considered in the manuscript. (A,B) BPI initialisation with A) volume fraction $\phi=1${\%} and B) $\phi=5${\%}. (C,D) ``quenched system'' with C) $\phi=1${\%} and D) $\phi=5${\%}.}
\end{figure}

\newpage

{\bf Supporting References}

@@

@@

@@

% MY NOTES

KEVIN

All particle radius $a = 2.3$.

Fluid properties: $\rho_0 = 1$; $\eta = 0.01$; $\zeta = 10\eta$;
temperature $T =2.133333 10^{-9}$.

Free energy:
$A_0 = 0.01$; $\gamma = 3.085143$; $q_0 = 0.09817477$ (and $p = 64$);
$\kappa_1 = \kappa_2 = 0.018971857$;
Chirality $\bar{\kappa} = 0.8$; $\tau = -0.25$

Amplitude order $ 0.0001$; Computed surface order $0.3509236$;
Rotational diffusion $\Gamma = 0.5$
Initialisation is 'random'.

Wall surface $W = 0.018971957$, ie., $W/\kappa$ = 1 length

Weak colloid anchoring $W = 0.0018971957$, ie., $W/\kappa = 0.1$ length
and $Wa/\kappa = 0.23$.

Strong colloid anchoring $W = 0.037943914$, ie., $W/\kappa = 2.0$ length
or $Wa/\kappa = 4.6$.

JUHO (BPI)

Fluid properties: $\rho_0 = 1$; $\eta = 0.01$; $\zeta = \eta$;
Temperature: $T = 2.133333 10^{-9}$.

Free energy: $A_0 = 0.01$; $\gamma = 3.0$; $q_0 = 0.13884$; (and $p = 45.25$
is $32\sqrt{2}$);
$\kappa_1 = \kappa_2 = 0.00706095$;
Chirality $\bar{\kappa} = 0.7$ Reduced temperature $\tau = 0.0$.

Amplitude of order $-0.2$; Computed surface order:
Rotational diffusion $\Gamma = 0.5$

Wall irrelevant

Weak colloid anchoring $W = 0.000706095$, ie., $W/\kappa = 0.1$ length
and $Wa / \kappa = 2.3$.

Strong colloid anchoring $W = 0.00706095$, i.e., $W/\kappa = 1.0$
and $Wa/\kappa = 1.0$.

JUHO (QUENCH)

Fluid properties: $\rho_0 = 1$; $\eta = 0.01$; $\zeta = \eta$;
Temperature: $T = 2.133333 10^{-9}$.

Free energy: $A_0 = 0.01$; $\gamma = 3.1$; $q_0 = 0.13884$; (and $p = 45.25$
is $32\sqrt{2}$);
$\kappa_1 = \kappa_2 = 0.00706095$;
Chirality $\bar{\kappa} = 0.68862$ Reduced temperature $\tau = -2.90322$.

>>>>

Explicitly, the free energy density is:
\begin{eqnarray}
f(Q_{\alpha\beta}) &=& {\textstyle \frac{1}{2}} A_0
 \left(1- {\textstyle \frac{1}{3}}\gamma \right)Q^2_{\alpha \beta}
-{\textstyle \frac{1}{3}}A_0\gamma Q_{\alpha \beta} Q_{\beta \gamma}Q_{\gamma \alpha} 
+ {\textstyle \frac{1}{4}} A_0\gamma (Q^2_{\alpha \beta})^2  \nonumber\\
&+& {\textstyle \frac{1}{2}} K(\varepsilon_{\alpha \gamma \delta}
\partial_\gamma Q_{\delta \beta} + 2 q_0 Q_{\alpha \beta})^2
+ {\textstyle \frac{1}{2}} K (\partial_\beta Q_{\alpha \beta})^2.
\label{free}
\end{eqnarray}
Here, repeated indices are summed over, while terms of the form
$Q^2_{\alpha\beta}$ should be expanded to read $Q_{\alpha\beta}Q_{\alpha\beta}$;
$\varepsilon_{\alpha\gamma\delta}$ is the permutation tensor.
The first three terms are a bulk free energy density whose overall scale
is set by $A_0$ (discussed further below); $\gamma$ is a control parameter
related to a reduced temperature. Varying the latter in the absence of
chiral terms ($q_0=0$) gives an isotropic-nematic transition at
$\gamma = 2.7$ with a mean-field spinodal instability at $\gamma = 3$.

The remaining two terms of the free energy density in Eq.~\ref{free} describe
distortions of the order parameter field. In theoretical work and when
describing a generic rather than a specific system, it is 
conventional~\cite{blue1,deGennes} to
assume that splay, bend and twist deformations of the director are equally
costly, that is, there is a single elastic constant $K$. The parameter $q_0$
is related to the helical pitch length $p$ via $q_0=2\pi/p$,
describing one full turn of the director in the cholesteric phase.

There are two dimensionless numbers, which are commonly referred to as
$\kappa$, the chirality, and  $\tau$, the reduced temperature
\cite{blue1} which can be used to characterise the system. In terms of
the parameters in the free energy density, these are:
\begin{eqnarray}\label{cntrl-param} 
\tau&=&\frac{27(1-\gamma/3)}{\gamma}\label{tau}\\
\kappa&=&\sqrt{\frac{108\ K\, q_0^2}{A_0\, \gamma}}\label{kappa}.
\end{eqnarray}
If the free energy density Eq.~\ref{free} is made dimensionless, $\tau$
appears as prefactor of the term  quadratic in $Q_{\alpha\beta}$,
whereas $\kappa$
quantifies the ratio between bulk and  gradient free energy terms. The
chirality and reduced temperature may be used to characterise an equilibrium
phase diagram of the blue phases in bulk cholesteric liquid crystals
\cite{blue1,oliver1}.

In this work, the chirality and reduced temperature are always set to lie
within the blue phase I region of the equilibrium phase diagram. EXCEPT?

@@

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_a}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
a further dimensionless number, 
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_a}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}

@@

The confining solid walls, which are flat and stationary, have the
same normal boundary condition as at colloid surface. In addition,
degenerate planar anchoring is applied at the walls, where the
preferred nematic director may adopt any orientation parallel
to the surface. The preferred direction is determined explicitly
by projecting the local fluid order parameter to the plane of the
wall \cite{fournier2005}. Eq.~\ref{eq:surf_q} and Eq.~\ref{eq:surf_q0}
are then used as before.

@@

The parameters for the free energy are chosen to be representative
of blue phase I: chirality $\kappa \sim 0.7$ and reduced temperature
$\tau = -0.25$.
The full free energy parameters are shown in Table~\ref{table:params}.
The order parameter tensor $Q_{\alpha\beta}$ for bulk blue phase I is
initialised from an approximation in high chirality limit
\cite{blue1,oliver1}. Systems of 128$^3$ lattice sites are
used with a pitch length of $p = 32\sqrt{2}$, which accommodates 4$^3$
unit cells with fully periodic boundary conditions.
Simulations at different solid volume fractions (1\% and 4\%), and different 
surface anchoring strengths ($Wa/K = 0.23$ and $Wa/K = 2.3$) representing
``weak'' and ``strong'' anchoring are run for two million simulation
time steps.
The initial distribution of the colloid positions is set at random
for the required solid volume fraction; the colloids are initially
at rest.

To generate a disordered network of disclination lines, a ``quench''
is performed in which the order parameter is initialised via a
locally chosen random director, and
Eq.~\ref{eq:surf_q} employed with a small amplitude $q^0 = 10^{-7}$ to
provide an order parameter.
We avoid the mean-field spinodal point at $\gamma = 3.0$:
the parameters are the same as for the initial simulations with
pitch $p = 32\sqrt{2}$. The chirality and reduced temperature (see
Table~\ref{table:params}) remain appropriate for blue phase I.
Colloids are added as before and the simulations run for 2 million
simulation steps.

@@

Here, a narrow sandwich of fluid is placed between flat walls in
perpendicular to the narrow coordinate direction, and with periodic
boundaries in the other two directions. The system size used in all
cases is 256$^2 \times$56 lattice sites. Each simulation is a ``quench''
in a similar manner to that used in the bulk: the order parameter is
initialised randomly. The free energy parameters (see Table~\ref{table:params})
are again appropriate for (equilibrium) blue phase I
(chirality $\kappa = 0.8$ and
reduced temperature $\tau = -0.25$). The cholesteric pitch is set
to be $p = 64$, providing a slight better resolution than the bulk
simulations.

The surface anchoring for colloids is as before: always normal, but with
$Wa/K = 0.23$ and $Wa/K = 2.3$. The colloid solid volume fractions are
1\% and 4\%. Wall anchoring is either normal or planar with $W/K = 1$
in simulation units. All simulations are run for two million simulation
time steps.

{\bf 4.3 External electric field}

OLIVER PLEASE ADD short description of protocol here. AND the
parameters in the table.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
         & $A_0$ & $\gamma$ & $K$ & $p$ & $\kappa$ & $\tau$ & $Wa/K$\\
\hline 
Bulk BPI & 0.01 & 3.086 & 0.007061 & 32$\sqrt{2}$ & 0.6902 & -0.2500 &
0.23, 2.3\\
\hline
Bulk Quench & 0.01 & 3.086 & 0.007061 & 32$\sqrt{2}$ & 0.6902 & -0.2500 &
0.23, 2.3 \\
\hline
Confined & 0.01 & 3.086 & 0.01897 & 64 & 0.8 & -0.2500 & 0.23, 2.3\\
\hline
External field & a0& gamma & K & p & kappa & tau & Wa/K\\ 
\hline
\end{tabular}
\end{center}
\caption{Free energy parameters used in the simulations. Bulk BPI
correspond to Fig.~1 (A-D), bulk quench to Fig.~1 (E--H); confined
geometries are shown in Fig.~2; and the external field case is
relevant for Fig.~3.}
\label{table:params}
\end{table}

In all cases visualisation of the defect lines is carried out via
an isosurface of the scalar order parameter $q$ determined from the
largest eigenvalue of the local order parameter tensor. A low value
of the scalar order parameter $q$ (typically in the range 0.12--0.14)
unambiguously identifies the disclinations.

@@

To get from simulation to physical units, we need to calibrate scales of
length, energy and time in the simulations. For similar mappings see,
e.g.~\cite{denniston2}.

The length scale can be set by mapping our half pitch to a realistic blue
phase unit cell size, which are in the 100-500 nm range \cite{blue1}. 
For bulk simulations (Fig.~1 and Fig.~3 in the main text), a suitable choice
is one where one simulation unit (lattice site) corresponds to 10~nm. Therefore
the colloidal size in Figs.~1 and~3 corresponds to $\sim$ 50~nm.

To obtain an energy scale, we choose $A_0 \simeq 10^6$~Pa, which is 
reasonable from Ref.~\cite{blue1} (this is also the choice in
Ref.~\cite{oliver2}). This choice leads to a simulation unit of
free energy (or stress) equal to $10^{8}$~Pa in Figs.~1 and~3.
From the energy and length scales one can map elastic constants to
physical units: for instance, the simulation in Fig.~1 corresponds to
a liquid crystal with splay, bend and twist (Frank) elastic constants
equal to 35 pN (or equivalently $K=$70~pN). That is,
\begin{equation}
\frac{a}{b^2} = \frac{27}{2 A_0\gamma} \sim 2-5 \times 10^{-6}
{\rm J}^{-1}\, {\rm m}^{3}.
\end{equation}
WHAT ARE a/b???

For the timescale calibration we use the formula (see e.g.~\cite{denniston})
\begin{equation}
\gamma_1=\frac{2q^2}{\Gamma}
\end{equation}
which relates the rotational viscosity $\gamma_1$, to the ordering strength
$q$ and the order parameter mobility $\Gamma$.
In our simulations, we choose $\Gamma = 0.5$, while the value of $\gamma$
used in Fig.~1 leads to $q=1/2$.   
For real liquid crystalline materials, $\gamma_1$ usually lies in range
$10^{-2}-1$ Pa s~\cite{deGennes}; we choose for definiteness
$\gamma_1 = 1$~Pa~s (equivalently 10 poise). Given the previous mapping for
free energy (or stress) units, this leads to a simulation unit of time equal
to $10^{-8}$~s. The total simulated time in Figs. 1-3 is therefore in the
millisecond range.

To calibrate the electric field strength 
we equate the value of the dimensionless parameter ${\cal E}$ in our 
simulations with that of a hypothetical experiment.
Given the previous value for $A_0$, and a dielectric anisotropy of 10
(or equivalently $\epsilon_a=10\epsilon_0$, with $\epsilon_0$ the
dielectric permittivity of vacuum), we then obtain that an electric field
of 5~V/$\mu$m corresponds to a value of ${\cal E}\sim$0.01.

Given these mappings, all quantities can be easily transformed from
simulation to physical units and vice versa. For example Figs.~1 and~2
use an isotropic viscosity $\eta \simeq$~0.01 and~0.1 in simulation units,
which translate into $0.01-0.1$~Pa~s respectively. This is sensible (if 
slightly low) for a molecular nematogen in the isotropic phase. 
The effective viscosity in ordered phases is higher, but this is ensured
in our model by the coupling to the order parameter.
We note that the density is set to unity in our simulations: this
corresponds to a fluid density which is about a thousand times larger
than in experiments. This however makes no difference in practice as
the Reynolds number $= \rho V\Lambda/\eta$ (where $V$ is a typical
velocity, in our simulations around $10^{-5}$ in simulation units) 
remains small enough~\cite{codef}.

To summarize, our simulations represent BP-forming materials, with the
interpretation of the simulation units for length, time, and energy
density being close to 10~nm, 10~ns, and 100~MPa respectively.

@@

@@

\vfill
\pagebreak

@@

\bibitem{codef} M. E. Cates,  et al., 
\textit{J. Phys. Condens. Matter} \textbf{16},  3903 (2004).

@@

\vfill\pagebreak

{\bf Supporting Figures and Animations}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.8]{doubletwist.pdf}
\end{center}
\caption{
(A) Schematic representation of the director field
in a cholesteric (left) and within a double twist cylinder (right).
(B) Snapshot of the disclination network in an equilibrated 
blue phase I structure. (C) Same as (B) for an amorphous blue phase III
network.}
\end{figure}

\vfill\pagebreak

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.35]{support-fig2.pdf}
\end{center}
\caption{As for main manuscript Fig.~1 A--D
(with liquid crystal order parameter
initialised to equilibrium blue phase I structure), but for the entire
simulation system of 128$^3$ lattice sites. (A)  solid volume fraction~1\%
and weak anchoring;  (B) 4\% solid volume fraction and weak anchoring;
(C) 1\% solid volume fraction and strong anchoring; (D) 4\% solid
volume fraction and strong anchoring. The view direction in the
main figure is from the left here. There is a reference
particle at the bottom left in each panel which does not take part
in the simulation.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.35]{support-fig3.pdf}
\end{center}
\caption{As for main manuscript Fig.~1 E--H (initialised via a ``quench''
to generate a disordered network), but for the entire simulation
system of 128$^3$ lattice sites. (E) solid volume fraction 1\% and
weak anchoring; (F) 4\% solid volume fraction and weak anchoring;
(G) 1\% solid volume fraction and strong anchoring; (H) 4\% solid
volume fraction and strong anchoring. Again, there is a reference
particle at the bottom left in each case which does not take part
in the simulation.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.42]{support-fig4.pdf}
\end{center}
\caption{As for main manuscript Fig.~2 A--D (confined geometry with normal
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (A) solid volume fraction 1\%
and weak anchoring; (B) 4\% solid volume fraction and weak anchoring;
(C) 1\% solid volume fraction and strong anchoring; (D) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.42]{support-fig5.pdf}
\end{center}
\caption{As for main manuscript Fig.~2 E--H (confined geometry with planar
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (E) colloid solid volume fraction 1\%
and weak anchoring; (F) 4\% solid volume fraction and weak anchoring;
(G) 1\% solid volume fraction and strong anchoring; (H) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\begin{figure}[t]
\begin{center}
%%\includegraphics[width=0.95\columnwidth]{s5.jpg}
\end{center}
\caption{Pair correlation function for weak ($WR/K=0.23$) and strong ($WR/K=2.3$) anchoring cases considered in the manuscript. (A,B) BPI initialisation with A) volume fraction $\phi=1${\%} and B) $\phi=5${\%}. (C,D) ``quenched system'' with C) $\phi=1${\%} and D) $\phi=5${\%}.}
\end{figure}

%{\bf Captions for Supplementary Movies}

@@
